An agent-based model for city networks based on interactions between firms

### 2020-03-04 main.tex
.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{mdframed}

@@

\author{J. Raimbault$^{1,2,3}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
}

@@

Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available). We then join this data with the GHS-FUA dataset for functional urban areas \cite{Florczyk2019ghs}, which are the relevant objects for applying our model. Starting from a firm-level dataset of 1,562,862 nodes and  1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links. Weight of links are obtained by computing the owned share of turnover at destination, i.e. $w_{ij} = \sum_{k \in i,l \in j} s_{kl} \cdot T_l$ where $s_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$. Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area. This quantity is highly correlated with area-level GDP included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$), so we expect to capture size effects while keeping the consistence of a single main data source. An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector. Detailed sectors of firms are available up to the fourth digit for the NACE classification in the raw data. We consider the first level classification (21 categories) and compute sector proportions accordingly.

@@

To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis which is similar to unconstrained spatial interaction models \cite{wilson1975some}.
% poisson \cite{flowerdew1988fitting}

@@

We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by
\begin{equation}
    s\left(S_{ik},S_{jk}\right) = \sum_{k=1}^{K} S_{ik} S_{jk}
\end{equation}

@@

\bibliographystyle{apalike}
%\bibliographystyle{unsrt}
\bibliography{biblio}

>>>>

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{natbib,graphicx,mdframed,hyperref,threeparttable}

@@

\author{J. Raimbault$^{1,2,3,\ast}$, N. Zdanowska$^{1,3}$ and E. Arcaute$^1$\\\medskip\small
$^{1}$ Centre for Advanced Spatial Analysis, UCL; $^{2}$ UPS CNRS 3611 ISC-PIF; $^{3}$ UMR CNRS 8504 G{\'e}ographie-cit{\'e}s\\
$^{\ast}$ \texttt{juste.raimbault@polytechnique.edu}
}

@@

Companies from the AMADEUS database are georeferenced using the Geonames database (using postcode or address depending on the information available). We then join this data with the GHS-FUA dataset for functional urban areas \cite{Florczyk2019ghs}, which are the relevant objects for applying our model. Starting from a firm-level dataset of 1,562,862 nodes and  1,866,936 links, we obtain a directed network of 573 urban areas and 9323 ownership links. Weight of links are obtained by computing the owned share of turnover at destination, i.e. $w_{ij} = \sum_{k \in i,l \in j} p_{kl} \cdot T_l$ where $p_{kl}$ of the ownership share of company $l$ by company $k$ and $T_l$ the turnover of company $l$. Node attributes include an economic size of urban areas, that we compute as the sum of turnovers of companies within the area. This quantity is highly correlated with area-level GDP included in the GHS database (Spearman correlation $\rho = 0.71$) and also with population ($\rho = 0.60$), so we expect to capture size effects while keeping the consistence of a single main data source. An other attribute of nodes is the industrial sector composition, expressed as a proportion of turnover in the area associated to a given industrial sector. Detailed sectors of firms are available up to the fourth digit for the NACE classification in the raw data. We consider the first level classification (21 categories) and compute sector proportions accordingly. This allows us to define an industrial proximity matrix between urban areas, taking a cosine similarity between vectors as $s_{ij} = \sum_{k=1}^{K} S_{ik}\cdot S_{jk}$ where $S_{ik}$ are sector proportions.

@@

To have a first insight into factors playing a role in the formation of links, we proceed to a statistical analysis which is similar to unconstrained spatial interaction models \cite{wilson1975some}. We test five different statistical models, including progressively the processes candidate to be included in the network generative model. The most general model is
\[
\log(w_{ij}) = \log(d_{ij}) + \log(T_i) + \log(T_j) + \log(s_{ij}) + \alpha_{c_i,c_j} + \varepsilon_{ij}
\]
Estimation results are given in Table~\ref{tab:reg}.
% poisson \cite{flowerdew1988fitting}

%%%%%%%%%%%%%
\begin{table}[h]
\caption{Ordinary Least Square estimations of statistical models explaining $\log(w_{ij})$ (standard errors in parenthesis). ***, ** and * respectively indicate 0.01, 0.05 and 0.1 levels of significance. For the fixed effect by origin-destination pairs of countries, we give the proportion of significant coefficients (where $p<0.1$). Models (1) to (4) are ordinary least square models, while model (5) is a Poisson generalized linear model with logarithmic link.\label{tab:reg}}
\medskip
\begin{tabular}{|l|c|c|c|c|c|}
\hline
Model  & (1) & (2) & (3) & (4) & (5) \\ 
\hline
$\log(d_{ij})$ &      -0.06** (0.03) &   -0.11** (0.05)  & -0.41*** (0.02)  & -0.35*** (0.04)  &  -0.26*** ($5\cdot 10^{-6}$) \\
$\log(T_i)$ &   &   & 0.56*** (0.01) &  0.56*** (0.01) & 0.79*** ($2\cdot 10^{-6}$) \\
$\log(T_j)$ &     &   & 0.39*** (0.01) &  0.39*** (0.01) & 0.66***  ($1.5\cdot 10^{-6}$) \\
$\log(s_{ij})$ &     &   &  &  0.19*** (0.07) & 0.53*** ($9\cdot 10^{-6}$)  \\
Countries &    &  28.3\% &   &  26.5\% & 100\% \\
\hline
$R^2$ &       0.00059   &  0.17 & 0.21 &  0.31  &  0.56 \\
AIC &        44304   &  43578  &  42131  & 41917 &  $4.41\cdot 10^{11}$ \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%

@@

%We take a cosine similarity between the sector structures, given with such probability vectors (size effect is already taken into account) by

@@

%\bibliographystyle{apalike}
\bibliographystyle{unsrt}
\bibliography{biblio}
